---
title: "TV_Rating_Viz"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
```
## Get the Data
```{r}
raw_data <- read.csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-01-08/IMDb_Economist_tv_ratings.csv")
summary(raw_data)
```
## Cleaning the data
We want to make sure that R understands the values in the columns. TO do this, we make sure that 
the dates are understood as dates, not characters.

```{r}
clean_data <- raw_data %>% mutate(date = ymd(date))

str(clean_data)
```
## Descriptive Statistics
Let's get to know the data a bit better with visualizations and counting.
```{r}
clean_data %>% ggplot(aes(x=log(share), y=av_rating)) + geom_point()
```

```{r}
clean_data %>% 
  group_by(titleId, title) %>% 
  summarise(number_of_seasons = max(seasonNumber)) %>% 
  arrange(desc(number_of_seasons))
```

## Extract new data from existing data
In this case, we want to know which series had higher ratings at the end of their run, 
than at the beginning.
In order to work that out, we need to know the rating of the first season and the rating of the 
lat season. Additionally, we want the number of seasons by counting by counting the occurences of
a series (because there might be some missing data).

```{r}
transformed_data <- clean_data %>%
  group_by(titleId, title) %>%
  summarise(first_season_rating = av_rating[which.min(seasonNumber)], 
            last_season_rating = av_rating[which.max(seasonNumber)],
            first_season_recorded = min(seasonNumber),
            last_season_recorded = max(seasonNumber),
            number_of_seasons = n()) %>%
  mutate(finished_with_higher_rating = first_season_rating < last_season_rating)
head(transformed_data)
```

Now let's count how many seasons had higher ratings before than after . . .
```{r}
 transformed_data %>% 
  group_by(finished_with_higher_rating) %>% 
  count()
```

```{r}
transformed_data %>%
  ggplot() +
  geom_point(aes(x=first_season_rating, 
                 y=last_season_rating,colour=finished_with_higher_rating))
```


But in this case we're not excluding data where there are only 1 season, which is skewing our data. We can see this from the straight line of dots (start rating is equal to the end rating).
So let's exclude those data using `filter`:
```{r}
 transformed_data %>% 
  filter(number_of_seasons > 1) %>% 
  group_by(finished_with_higher_rating) %>% 
  count()
```

```{r}
transformed_data %>%
  filter(number_of_seasons > 1) %>%
  ggplot() +
    geom_point(aes(x=first_season_rating, y=last_season_rating, colour=finished_with_higher_rating), alpha=0.5) +
    geom_abline(intercept = 0, slope = 1)

```
From our rudimentary analysis, it would appear as though more series end with a higher rating than they started.

## Further Analysis
What movies ended with much worse ratings than when they began?
```{r}
worst_performing_series <- transformed_data %>%
  filter(number_of_seasons > 1) %>%
  group_by(titleId,title) %>%
  summarize(drop_in_ratings = last_season_rating - first_season_rating) %>%
  head(20)

worst_performing_series
```
